---
title: "Homework 1 -- Due Monday September 21, 2015"
author: "Cheryl Calhoun"
date: "09/03/2015"
output:
  pdf_document: default
  html_document: default
---

Add your code blocks to this document and alter the Author name to yours. Your submission will consist of your own R Markdown file plus the compiled HTML version of the document.

I encourage you to submit your completed version as soon as possible. We will grade this quickly so that any opportunities for a regrade can be done quickly.

## Assignment

This homework will continue using `dplyr` and the operations to join related data frames together. We will use the `GSS-extract-2.RData` data set from class on September 14.

Recall the in-class exercise from Monday the 14th wherein we made smaller data frames based on two factor variables, first by making a long table. Let's make a few more.

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
## Seting up the session.
## Load dplyr.
library(dplyr)

## Set working directory.
## setwd("C:/Users/Cheryl/OneDrive/Education/UF/2015/Fall/EDF6938/Week 3")
## setwd("C:/Users/07001412/OneDrive/Education/UF/2015/Fall/EDF6938/Week 3")

## Load data file.
load ("GSS-extract-2.RData")

## Rename variable for same-sex marriage
subGSS.2 <- rename(subGSS.2, SSMar = marhomo)
```

## Part I

1. Construct a long data frame wherein you find the mean hours of TV watched and radio listened to for each subgroup of marital status and education achieved, only with respondents after 1980 who are under 50 years of age. Join this to the original data frame.

```{r, eval=TRUE}

## Filter out records which are missing values. Create subGSS.2I1 to hold data set for homework 2, PartI, Question 1. 
subGSS.2I1 <- filter(subGSS.2, !is.na(radiohrs) & !is.na(tvhours))

## Filter out respondents over 50 years, and before 1980
subGSS.2I1 <- filter(subGSS.2I1, age < 50 & year >= 1980)

## Group data set by marital status and degree level, calculate mean tvhours for each level, calculate mean radiohours for each level
by_age_year <- subGSS.2I1 %>% 
  group_by (degree, marital) %>% 
  mutate (MTVhours = mean(tvhours, na.rm=TRUE), Mradiohrs = mean(radiohrs, na.rm=TRUE))

## display by_age_year
by_age_year

## Add MTVhours & Mradiohrs back into subGSS.2I1
subGSS.2I1 <- left_join(subGSS.2, by_age_year)

## Verify MTVhours and Mradiohrs are joined with subGSS.2.
head(subGSS.2I1, 2)
```

**Results:** subGSS.2I1 now contains 13 variables (added mean values for radio hours and TV hours), vs. the 11 variables it started with. The data set represents responses obtained after 1980 for respondents who are under 50 years of age. 

2. Create a scatterplot using the `plot` function of TV hours watched (y) versus income (x) for all individuals. Add arguments `main`, `xlab` and `ylab` for a title and axis labels. 

```{r, eval=TRUE}

## Create a scatterplot of TV hours watched (y) versus income (x) for all individuals in subGSS.2I2. 

## Remove rows with missing values for conrinc and assign to subGSS.2I2.
subGSS.2I2 <- filter(subGSS.2I1, !is.na(conrinc))

plot(subGSS.2I2$conrinc, subGSS.2I2$tvhours, type="p", main="TV hours watched based on income levels", sub = "Responses collected after 1980, for respondents under 50 years of age", xlab="Income", ylab="TV Hours Watched")

```

**Results: ** The resulting plot shows TV hours watched to income for responses collected after 1980 for respondents who are under age 50. 

3. Use the function `points`, which takes the same first two arguments as `plot` to add points to this scatterplot corresponding to the group mean for TV hours watched (y) versus individual income (x). Add the option `co="red"` to this command so that the "group mean" points are clearly visible.

```{r, eval=TRUE}

## Create a scatterplot of TV hours watched (y) versus income (x) for all individuals in subGSS.2I2. 

plot(subGSS.2I2$conrinc, subGSS.2I2$tvhours, type="p", main="TV hours watched based on income levels  with Mean TV hours in red", sub = "Responses collected after 1980, for respondents under 50 years of age", xlab="Income", ylab="TV Hours Watched")

## Adding mean (x,y) value in solid red.
points(subGSS.2I2$conrinc, subGSS.2I2$MTVhours, pch = 19, col="red")
```

**Results: ** The resulting plot shows TV hours watched to income for responses collected after 1980 for respondents who are under age 50. Overlayed on the original plot, is a plot of income to mean TV hours for each category (year, marital, degree) shown in red dots.

4. Construct a long data frame wherein you find the mean hours of TV watched and radio listened to for each subgroup of opinion of marijuana legalization and education achieved, only with respondents who make less than $50,000. Join this to the original data frame.

```{r, eval=TRUE}
## Filter out records which are missing values for radiohrs and tvhours. Create subGSS.2I4 to hold data set for Question 4.
subGSS.2I4 <- filter(subGSS.2, !is.na(radiohrs) & !is.na(tvhours))

## `filter` out respondents over 50 year.
subGSS.2I4 <- filter(subGSS.2I4, conrinc < 50000)

## Group data set by degree and legalization opinion, calculate mean tvhours for each level, calculate mean radiohours for each level
by_degree_grass <- subGSS.2I4 %>% 
  group_by (degree, grass) %>% 
  mutate (MTVhours = mean(tvhours, na.rm=TRUE), Mradiohrs = mean(radiohrs, na.rm=TRUE))

## display by_degree_grass
by_degree_grass

## Add MTVhours & Mradiohrs back into subGSS.2I4
subGSS.2I4 <- left_join(subGSS.2, by_degree_grass)
str(subGSS.2I4)
```

**Results:** The resulting plot shows TV hours watched to income for responses collected after 1980 for respondents who are under age 50.

5. Create a scatterplot using the `plot` function of TV hours watched (y) versus income (x) for all individuals. Add arguments `main`, `xlab` and `ylab` for a title and axis labels. 

```{r, eval=TRUE}

## Create a scatterplot of TV hours watched (y) versus income (x) for all individuals in subGSS.2I2. 

## Remove rows with missing values for conrinc and assign to subGSS.2I2.
subGSS.2I4 <- filter(subGSS.2I4, !is.na(conrinc))

plot(subGSS.2I4$conrinc, subGSS.2I4$tvhours, type="p", main="TV hours watched based on income levels", sub = "Responses for respondents under 50 years of age", xlab="Income", ylab="TV Hours Watched")

```

**Results: ** The resulting plot shows TV hours watched to income for respondents who are under age 50. 

6. Use the function `points`, which takes the same first two arguments as `plot` to add points to this scatterplot corresponding to the group mean for TV hours watched (y) versus individual income (x). Add the option `col="red"` to this command so that the "group mean" points are clearly visible.

```{r, eval=TRUE}

## Create a scatterplot of TV hours watched (y) versus income (x) for all individuals in subGSS.2I2. 

plot(subGSS.2I4$conrinc, subGSS.2I4$tvhours, type="p", main="TV hours watched based on income levels, Mean TV hours in red", sub = "Responses for respondents under 50 years of age", xlab="Income", ylab="TV Hours Watched")


## Adding mean (x,y) value in solid red.
points(subGSS.2I4$conrinc, subGSS.2I4$MTVhours, pch = 19, col="red")

```

**Results: ** The resulting plot shows TV hours watched to income for respondents who are under age 50. Overlayed on the original plot, is a plot of income to mean TV hours for each category (year, marital, degree) shown in red dots.

## Part II

We have class records for several different groups of people. There is an overlap among the rosters of the three classes and we wish to compose a complete grade sheet for all students. You will need to load the file `hw2students.RData` into your workspace as part of your script.

7. Join the three tables together into one. Which `join` function will you need to use to include all student names across all three tables?

```{r, eval=TRUE}
## Load dplyr.
library(dplyr)
## Set working directory.
##setwd("C:/Users/Cheryl/OneDrive/Education/UF/2015/Fall/EDF6938/Week 3")
## Load data file.
load ("hw2students.RData")
## Join report.card.1, report.card.2, and report.card.3 together into report.card. Use full_join to retain all student records.
report.card <- full_join(report.card.1, report.card.2, by="Name")
report.card <- full_join(report.card, report.card.3, by="Name")
```

**Results:** The resulting report.card data frame contains all students with grades from all 9 courses as applicable. 

8. Determine the mean grade achieved for all students in the classes they take. You will need to use `group_by`, `mutate` and `mean` while using a function to remove the missing values for each row. Add this to the data frame

```{r, eval=TRUE}

## Using pipe & group_by to create mean for each row. 
report.card.avg <- report.card %>% group_by(Name) %>% mutate(Avg = mean(c(Math, English, History, Biology, Chemistry, Physics, Semiotics, Mixology, Showmanship), na.rm=TRUE))
## Display report.card.avg
report.card.avg
str(report.card.avg, list.len = 12)
head(report.card.avg$Avg)
```

**Results:** The resulting report.card.avg contains a new column (Avg) with the average grade from all courses taken.

9. Determine the number of classes taken by each student and add this to the data frame. Find the top-achieving student in each group: those who took all 9 classes, only 6, and only 3.

```{r, eval=TRUE}

## Using pipe & group_by to create count for each row. 
ungroup(report.card.avg)
report.card.classes <- 
  report.card.avg %>% 
  mutate(Classes = (sum(c(Math, English, History, Biology, Chemistry, Physics, Semiotics, Mixology, Showmanship), na.rm=TRUE))/Avg)
top_n(group_by(report.card.classes, Classes), 1, Avg)
str(report.card.avg, list.len = 12)

```

**Results:** The results produce a list of the student with the highest average grade from each level of participation (3 courses, 6 courses, 9 courses).
